### 2014-09-16 fundamentals/impl.tex
notes on discrete resources and materials and tracking

Discrete facility and material tracking is more realistic an dmore

>>>>

TODO: add blurb about discrete agent/facility tracking.

In Cyclus, agents can transfer discrete resource objects between each other.
Cyclus supports two types of resources:

\begin{itemize}

  \item Materials: These represent typical nuclear materials with
      nuclide-based compositions.

  \item Products: These can represent any user-defined thing - e.g. carbon
      credits, build permits, employees, etc.

\end{itemize}

All operations performed on every resource object (splitting, combining,
decay, etc.) are tracked in detail.  This information includes the creating
agent for when resources are newly created and introduced into the simulation.
The parentage of each resource is also tracked. This makes it possible to
follow the travel history of resource objects as they are transferred between
agents.

In large simulations there may be many material objects changing frequently.
Material decay can also contribute significantly to such changes.  In order to
help avoid unnecessary runtime performance and database space impacts,
compositions in Cyclus have some special features.  Compositions in Cyclus are
immutable.  This immutability allows multiple material objects to all hold
references to the same composition object safely.  When operations are
performed on a material that don't change the composition, any new resulting
materials hold a reference to the same, originating composition object.
Although the decay operation generates a new, separate composition (preserving
immutability), a reference to the new composition is placed in a cache shared
by the new an originating compositions. The next time the originating
composition is decayed, this cache will be checked for a prior calculation
with the same cumulative decay time.  If one exists, the cached composition is
used.  This cache is shared by all compositions that transitively come from
the same origin.  Composition immutability in concert with decay history
caching help eliminate many redundant decay calculations in addition to
reducing the total number of composition objects.  Since each composition
object is only recorded in the database once, significant space savings also
occurs.
